Title here

### 2020-08-06 analysis/paper/paper.Rmd
resultados

```{r asymsam-timeseries, fig.cap = "Time series for the asymmetric SAM and symmetric SAM.", fig.width=6, fig.height=3, fig.env="figure*"}
indices %>% 
  .[lev %in% c(30, 700) & term != "full"] %>% 
  # .[, estimate_norm := estimate_norm/sd(estimate_norm), by = .(lev, term)] %>% 
  ggplot(aes(time, estimate_norm)) +
  geom_line(aes(color = term)) +
  scale_color_brewer(NULL, palette = "Set1", labels = lab_sam) +
  scale_y_continuous(NULL) +
  scale_x_datetime(NULL, expand = c(0, 0)) +
  facet_grid(lev~., labeller = labeller(lev = lev.lab), scales = "free") +
  tag_facets()
```

Figure \ref{fig:asymsam-timeseries} shows the resulting Asymmetric and Symmetric time series corresponding to 700 and 30hPa. blablababla #FIXME

* stratosphere clearly nor normally distributed. a lot of values near 0 and some relatively high outliers. Especially true int he case of the asymmetric index. High frequency variablity.

* In both levels, there's correlation between the series (expected),

@@

Correlations between the Asymmetric and Symmetric series are rather constant throught the troposphere, fluctiating between `r signif(cors[lev >= 100, min(V1)], 2)` and `r signif(cors[lev >= 100, max(V1)], 2)` (Figure \ref{fig:cor-lev}). Futhermore, the cross-correlation of each series across levels --shown in Figure \ref{fig:cross-correlation}-- are high in the trosposphere (greater than 0.9) for both indices. This suggests that both the Asymmetric and the Symemtric component of the tropospheric SAM are highly vertically coherent, both in their individual evolution and their temporal relationship. This is to be expected since the SAM is mostly equivalent barotropic (citaaaa).

@@

Figure \ref{fig:cross-correlation}a) show the cross-correlation across levels for the Full SAM index. #FIXME

@@

@@

```{r ccf-levels, fig.cap = "Cross-correlation functions for each index and two differnet base levels."}
ccf_dt <- function(...) {
  a <- ccf(..., plot = FALSE)
  list(acf = as.vector(a$acf),
       lag = as.vector(a$lag))
}

@@

```{r compute-spectrums}
indices %>% 
  .[lev %in% c(50, 700)] %>% 
  # .[, estimate_norm := estimate_norm/sd(estimate_norm), by = .(lev, term)] %>% 
  .[, fftspectrum(estimate, 5, B = 5000, probs = c(0.025, 0.975)), by = .(term, lev)] -> spec
```

```{r A2, fig.cap = "Fourier spectrum of each timeseries. The shading indicates de 95\\% area derived by fitting an AR process to each series and bootstrapping 5000 simulated samples.", appendix = TRUE}
spec %>%
  .[term != "full"] %>% 
  ggplot(aes(1/freq/12, spec)) +
  geom_ribbon(aes(fill = term, ymin = `2%`, ymax = `98%`), position = "dodge", alpha = 0.1) +
  geom_line(aes(color = term, y = ar_spectrum), alpha = 0.15) +
  geom_line(aes(color = term)) +
  annotation_logticks(sides = "b") +
  scale_color_brewer(NULL, palette = "Set1", aesthetics = c("color", "fill")) +
  scale_x_log10("Period (years)", sec.axis = sec_axis(~.*12)) +
  facet_grid(lev ~ term,labeller = labeller(lev = lev.lab, term = lab_sam), scales = "free") +
  tag_facets() 
```

```{r A3, fig.cap = "Autocorrelation functions of each timeseries", appendix = TRUE}
acf_dt <- function(...) {
  a <- acf(..., plot = FALSE)
  list(acf = as.vector(a$acf),
       lag = as.vector(a$lag))
}
indices %>% 
  .[lev %in% c(30, 700)] %>% 
  .[, acf_dt(estimate_norm), by = .(lev, term)] %>% 
  ggplot(aes(lag, acf)) +
  geom_col(aes(fill = term)) +
  scale_fill_brewer(palette = "Set1") +
  facet_grid(term ~ lev,labeller = labeller(lev = lev.lab, term = lab_sam)) +
  tag_facets(position = "tr") 
```

@@

pp_regr_season %>% 
  filter_sa() %>% 
  .[, lon := ConvertLongitude(lon)] %>% 
  ggplot(aes(lon, lat)) +
  # geom_raster(aes(fill = estimate)) +
  geom_contour_fill(aes(z = estimate), na.fill = 0, breaks = AnchorBreaks(0, 0.25, 0)) +
  geom_contour2(aes(z = p.value), breaks = 0.01) +
  geom_contour_fill(aes(z = as.numeric(is.na(estimate))), breaks = 1) +
  stat_subset(aes(subset = p.value <= 0.01 & is.cross(lon, lat)), size = 0.4, alpha = 0.3) +
  geom_qmap(~filter_sa(.x)[, lon := ConvertLongitude(lon)], keep = 0.02) +
  scale_x_longitude() +
  scale_y_latitude() +
  scale_fill_divergent("Precipitation", guide = guide_colorstrip_bottom(), 
                       breaks = AnchorBreaks(0, 0.25, 0)) +
  # coord_polar() +
  ggalt::coord_proj("+proj=moll +lon_0=-60") +
  facet_grid(season~term, labeller = labeller(term = lab_sam)) +
  tag_facets() +
  axis_labs_smol 
```

```{r A6, appendix = TRUE, fig.cap = "Regressions maps resulting from performing one multiple regression (a. and b.) and from performing two simple regressions (c. and d.)"}
regr_dif <- hgt[lev %in% 700] %>% 
  indices_wide[., on = .NATURAL, allow.cartesian = TRUE] %>% 
  .[,  rbind(idcol = "regression", multiple = as.data.table(FitLm(hgt_a, sym, asym)),
             single = rbind(as.data.table(FitLm(hgt_a, sym)),
                            as.data.table(FitLm(hgt_a, asym)))),
    by = .(lon, lat)]

regr_dif %>% 
  rm_intercept() %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, term := factor(term, levels = names(lab_sam), ordered = TRUE)] %>% 
  .[lat <= -20] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = ZeroBreaks, global.breaks = FALSE) +
  # geom_contour2(aes(z = estimate, linetype = factor(-sign(..level..))), 
  # size = 0.3,
  # global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = ZeroBreaks,
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  # geom_contour2(data = ~.x[dataset == "NCEP"], aes(z = V1)) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20), labels = NULL) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-85, 85),
                       oob = scales::squish,
                       breaks = MakeBreaks(binwidth = 20, exclude = 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ regression, labeller = labeller(term = lab_sam, lev = lev.lab)) +
  tag_facets() +
  axis_labs_smol 
```

```{r A7, appendix = TRUE, fig.cap = " "}
regr_dif %>% 
  rm_intercept() %>% 
  .[lat < -20] %>%
  # .[lev %in% c(50, 700)] %>% 
  .[, FitWave(estimate, 0:3), by = .(lat, regression, term)] %>% 
  ggplot(aes(lat, amplitude)) +
  geom_line(aes(color = factor(k))) +
  scale_color_brewer("Wave-number", palette = "Set1") +
  scale_y_continuous("Amplitude") +
  scale_x_latitude(ticks = 15) +
  coord_flip() +
  facet_grid(term~regression, labeller = labeller(term = lab_sam, lev = lev.lab), scales = "free_x") +
  tag_facets(position = "tr")
```

>>>>

```{r asymsam-timeseries, fig.cap = "Time series for the asymmetric SAM and symmetric SAM and density estimates.", fig.width=6, fig.height=3, fig.env="figure*"}
indices %>% 
  .[lev %in% c(30, 700) & term != "full"] %>% 
  # .[, estimate_norm := estimate_norm/sd(estimate_norm), by = .(lev, term)] %>% 
  ggplot(aes(time, estimate_norm)) +
  geom_line(aes(color = term)) +
  ggridges::geom_vridgeline(stat="ydensity", scale = -4e8, fill = NA,
                            aes(color = term, width = ..density.., x = min(time - 1e7))) +
  scale_color_brewer(NULL, palette = "Set1", labels = lab_sam) +
  scale_y_continuous(NULL) +
  # scale_x_datetime(NULL, expand = c(0, 0)) +
  facet_grid(lev~., labeller = labeller(lev = lev.lab), scales = "free") +
  tag_facets()
```

Figure \ref{fig:asymsam-timeseries} shows the resulting Asymmetric and Symmetric time series corresponding to 700 and 30hPa.  #FIXME

At first glance the series can be distinguished by their distributions. Whereas the tropospheric incides are aproximately normally distributed, the stratospheric indices are more long-tailed; that is, extreme values (both negative and positive) abound. The Asymmetric series have both more variability in the higher frequencies than the Symmetric series.

The stratospheric Symmetric SAM varies strongly with  a two-year period, which can be seen using spectral methods (Figure A3) or in the autocorrelation structure (Figure A4). There is a local peak at 2 years in the periodigram of the tropospheric Symmetric SAM also, although it's not statistically significant. In the troposphere the most significant peak of variability is found in the Asymmetric index at around `r 12*.3` months.

@@

Correlations between the Asymmetric and Symmetric series are constant throught the troposphere, fluctiating between `r signif(cors[lev >= 100, min(V1)], 2)` and `r signif(cors[lev >= 100, max(V1)], 2)` (Figure \ref{fig:cor-lev}). Futhermore, the cross-correlation of each series across levels --shown in Figure \ref{fig:cross-correlation}-- are high in the trosposphere (greater than 0.9) for both indices. This suggests that both the Asymmetric and the Symemtric component of the tropospheric SAM are highly vertically coherent, both in their individual evolution and their temporal relationship. This is to be expected since the SAM is mostly equivalent barotropic (citaaaa).

@@

Figure \ref{fig:cross-correlation}a) show the cross-correlation across levels for the Full, Symmetric and Asymmetric SAM indices. The high values below 100 hPa in panel reflects the vertical (zero-lag) coherency of the tropospheric SAM. Above 100 hPa correlation between levels falls of rapidly, indicating less coherent (zero-lag) variability. At the same time, there is a non negligible correlation between the troposphere and the lower-to-middle stratosphere. By examining panels b and c we see that the Asymemtric and Symmetric SAM  share the smae high level of coherency in the tropospherem but that the Asymmetric SAM is relatively more coherent in the stratosphere. The tropospheric-stratospheric connection is only present in the Symmetric SAM.

@@

??

@@

```{r A2 ccf-levels, fig.cap = "Cross-correlation functions for each index and two differnet base levels."}
ccf_dt <- function(...) {
  a <- ccf(..., plot = FALSE)
  list(acf = as.vector(a$acf),
       lag = as.vector(a$lag))
}

@@

```{r compute-spectrums}
indices %>% 
  .[lev %in% c(50, 700)] %>% 
  # .[, estimate_norm := estimate_norm/sd(estimate_norm), by = .(lev, term)] %>% 
  .[, fftspectrum(estimate, c(3, 5), B = 5000, probs = c(0.025, 0.975)), by = .(term, lev)] -> spec
```

```{r A3, fig.cap = "Fourier spectrum of each timeseries. The shading indicates de 95\\% area derived by fitting an AR process to each series and bootstrapping 5000 simulated samples.", appendix = TRUE}
spec %>%
  # .[term != "full"] %>% 
  ggplot(aes(1/freq/12, spec)) +
  geom_ribbon(aes(fill = term, ymin = `2%`, ymax = `98%`), position = "dodge", alpha = 0.1) +
  geom_line(aes(color = term, y = ar_spectrum), alpha = 0.15) +
  geom_line(aes(color = term)) +
  annotation_logticks(sides = "b") +
  scale_color_brewer(NULL, palette = "Set1", aesthetics = c("color", "fill")) +
  scale_x_log10("Period (years)", sec.axis = sec_axis(~.*12)) +
  facet_grid(lev ~ term,labeller = labeller(lev = lev.lab, term = lab_sam), scales = "free") +
  tag_facets() 
```

```{r A4, fig.cap = "Autocorrelation functions of each timeseries", appendix = TRUE}
acf_dt <- function(...) {
  a <- acf(..., plot = FALSE)
  list(acf = as.vector(a$acf),
       lag = as.vector(a$lag))
}
indices %>% 
  .[lev %in% c(30, 700)] %>% 
  .[, acf_dt(estimate_norm), by = .(lev, term)] %>% 
  ggplot(aes(lag, acf)) +
  geom_col(aes(fill = term)) +
  scale_fill_brewer(palette = "Set1") +
  facet_grid(term ~ lev,labeller = labeller(lev = lev.lab, term = lab_sam)) +
  tag_facets(position = "tr") 
```

@@

pp_regr_season %>% 
  filter_sa() %>% 
  .[, lon := ConvertLongitude(lon)] %>% 
  .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(season, term)] %>% 
  ggplot(aes(lon, lat)) +
  # geom_raster(aes(fill = estimate)) +
  geom_contour_fill(aes(z = estimate), na.fill = 0, breaks = AnchorBreaks(0, 0.25, 0)) +
  geom_contour2(aes(z = p.value), breaks = 0.01) +
  geom_contour_fill(aes(z = as.numeric(is.na(estimate))), breaks = 1) +
  stat_subset(aes(subset = p.value <= 0.01 & is.cross(lon, lat)), size = 0.4, alpha = 0.3) +
  geom_qmap(~filter_sa(.x)[, lon := ConvertLongitude(lon)], keep = 0.02) +
  scale_x_longitude() +
  scale_y_latitude() +
  scale_fill_divergent("Precipitation", guide = guide_colorstrip_bottom(), 
                       breaks = AnchorBreaks(0, 0.25, 0)) +
  # coord_polar() +
  # ggalt::coord_proj("+proj=moll +lon_0=-60") +
  facet_grid(season~term, labeller = labeller(term = lab_sam)) +
  tag_facets() +
  axis_labs_smol 
```

```{r A6, appendix = TRUE, fig.cap = "Regressions maps resulting from performing one multiple regression (column a) and from performing two simple regressions (column b)"}
regr_dif <- hgt[lev %in% 700] %>% 
  indices_wide[., on = .NATURAL, allow.cartesian = TRUE] %>% 
  .[,  rbind(idcol = "regression", multiple = as.data.table(FitLm(hgt_a, sym, asym)),
             single = rbind(as.data.table(FitLm(hgt_a, sym)),
                            as.data.table(FitLm(hgt_a, asym)))),
    by = .(lon, lat)]

regr_dif %>% 
  rm_intercept() %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, term := factor(term, levels = names(lab_sam), ordered = TRUE)] %>% 
  .[lat <= -20] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = ZeroBreaks, global.breaks = FALSE) +
  # geom_contour2(aes(z = estimate, linetype = factor(-sign(..level..))), 
  # size = 0.3,
  # global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = ZeroBreaks,
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  # geom_contour2(data = ~.x[dataset == "NCEP"], aes(z = V1)) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20), labels = NULL) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-85, 85),
                       oob = scales::squish,
                       breaks = MakeBreaks(binwidth = 20, exclude = 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ regression, labeller = labeller(term = lab_sam, lev = lev.lab)) +
  tag_facets("cr") +
  axis_labs_smol 
```

```{r A7, appendix = TRUE, fig.cap = "Zonal waves derives from the regression maps from performing one multiple regression (column a) and from performing two simple regressions (column b)"}
regr_dif %>% 
  rm_intercept() %>% 
  .[lat < -20] %>%
  # .[lev %in% c(50, 700)] %>% 
  .[, FitWave(estimate, 0:3), by = .(lat, regression, term)] %>% 
  ggplot(aes(lat, amplitude)) +
  geom_line(aes(color = factor(k))) +
  scale_color_brewer("Wave-number", palette = "Set1") +
  scale_y_continuous("Amplitude") +
  scale_x_latitude(ticks = 15) +
  coord_flip() +
  facet_grid(term~regression, labeller = labeller(term = lab_sam, lev = lev.lab), scales = "free_x") +
  tag_facets("cr", position = "tr")
```

### 2020-08-06 analysis/paper/paper.tex
render figuras
resultados

\begin{figure*}
\includegraphics{asymsam-timeseries-1} \caption[Time series for the asymmetric SAM and symmetric SAM]{Time series for the asymmetric SAM and symmetric SAM.}\label{fig:asymsam-timeseries}
\end{figure*}

Figure \ref{fig:asymsam-timeseries} shows the resulting Asymmetric and
Symmetric time series corresponding to 700 and 30hPa. blablababla
\#FIXME

\begin{itemize}
\item
  stratosphere clearly nor normally distributed. a lot of values near 0
  and some relatively high outliers. Especially true int he case of the
  asymmetric index. High frequency variablity.
\item
  In both levels, there's correlation between the series (expected),
\end{itemize}

Correlations between the Asymmetric and Symmetric series are rather
constant throught the troposphere, fluctiating between 0.39 and 0.45
(Figure \ref{fig:cor-lev}). Futhermore, the cross-correlation of each
series across levels --shown in Figure \ref{fig:cross-correlation}-- are
high in the trosposphere (greater than 0.9) for both indices. This
suggests that both the Asymmetric and the Symemtric component of the
tropospheric SAM are highly vertically coherent, both in their
individual evolution and their temporal relationship. This is to be
expected since the SAM is mostly equivalent barotropic (citaaaa).

@@

Figure \ref{fig:cross-correlation}a) show the cross-correlation across
levels for the Full SAM index. \#FIXME

@@

\begin{figure}
\includegraphics{ccf-levels-1} \caption[Cross-correlation functions for each index and two differnet base levels]{Cross-correlation functions for each index and two differnet base levels.}\label{fig:ccf-levels}
\end{figure}

\begin{figure}
\includegraphics{A2-1} \appendcaption{A2}{Fourier spectrum of each timeseries. The shading indicates de 95\% area derived by fitting an AR process to each series and bootstrapping 5000 simulated samples.}\label{fig:A2}
\end{figure}

\begin{figure}
\includegraphics{A3-1} \appendcaption{A3}{Autocorrelation functions of each timeseries}\label{fig:A3}
\end{figure}

@@

\begin{figure}
\includegraphics{A6-1} \appendcaption{A6}{Regressions maps resulting from performing one multiple regression (a. and b.) and from performing two simple regressions (c. and d.)}\label{fig:A6}
\end{figure}

\begin{figure}
\includegraphics{A7-1} \caption[ ]{ }\label{fig:A7}
\end{figure}

>>>>

\begin{figure*}
\includegraphics{asymsam-timeseries-1} \caption[Time series for the asymmetric SAM and symmetric SAM and density estimates]{Time series for the asymmetric SAM and symmetric SAM and density estimates.}\label{fig:asymsam-timeseries}
\end{figure*}

Figure \ref{fig:asymsam-timeseries} shows the resulting Asymmetric and
Symmetric time series corresponding to 700 and 30hPa. \#FIXME

At first glance the series can be distinguished by their distributions.
Whereas the tropospheric incides are aproximately normally distributed,
the stratospheric indices are more long-tailed; that is, extreme values
(both negative and positive) abound. The Asymmetric series have both
more variability in the higher frequencies than the Symmetric series.

The stratospheric Symmetric SAM varies strongly with a two-year period,
which can be seen using spectral methods (Figure A3) or in the
autocorrelation structure (Figure A4). There is a local peak at 2 years
in the periodigram of the tropospheric Symmetric SAM also, although it's
not statistically significant. In the troposphere the most significant
peak of variability is found in the Asymmetric index at around 3.6
months.

Correlations between the Asymmetric and Symmetric series are constant
throught the troposphere, fluctiating between 0.39 and 0.45 (Figure
\ref{fig:cor-lev}). Futhermore, the cross-correlation of each series
across levels --shown in Figure \ref{fig:cross-correlation}-- are high
in the trosposphere (greater than 0.9) for both indices. This suggests
that both the Asymmetric and the Symemtric component of the tropospheric
SAM are highly vertically coherent, both in their individual evolution
and their temporal relationship. This is to be expected since the SAM is
mostly equivalent barotropic (citaaaa).

@@

Figure \ref{fig:cross-correlation}a) show the cross-correlation across
levels for the Full, Symmetric and Asymmetric SAM indices. The high
values below 100 hPa in panel reflects the vertical (zero-lag) coherency
of the tropospheric SAM. Above 100 hPa correlation between levels falls
of rapidly, indicating less coherent (zero-lag) variability. At the same
time, there is a non negligible correlation between the troposphere and
the lower-to-middle stratosphere. By examining panels b and c we see
that the Asymemtric and Symmetric SAM share the smae high level of
coherency in the tropospherem but that the Asymmetric SAM is relatively
more coherent in the stratosphere. The tropospheric-stratospheric
connection is only present in the Symmetric SAM.

@@

\begin{figure}
\includegraphics{A2 ccf-levels-1} \caption[Cross-correlation functions for each index and two differnet base levels]{Cross-correlation functions for each index and two differnet base levels.}\label{fig:A2 ccf-levels}
\end{figure}

\begin{figure}
\includegraphics{A3-1} \appendcaption{A3}{Fourier spectrum of each timeseries. The shading indicates de 95\% area derived by fitting an AR process to each series and bootstrapping 5000 simulated samples.}\label{fig:A3}
\end{figure}

\begin{figure}
\includegraphics{A4-1} \appendcaption{A4}{Autocorrelation functions of each timeseries}\label{fig:A4}
\end{figure}

@@

\begin{figure}
\includegraphics{A6-1} \appendcaption{A6}{Regressions maps resulting from performing one multiple regression (column a) and from performing two simple regressions (column b)}\label{fig:A6}
\end{figure}

\begin{figure}
\includegraphics{A7-1} \appendcaption{A7}{Zonal waves derives from the regression maps from performing one multiple regression (column a) and from performing two simple regressions (column b)}\label{fig:A7}
\end{figure}

\begin{figure}
\includegraphics{A8-1} \appendcaption{A8}{Decadal trends of SAM indices for each season. }\label{fig:A8}
\end{figure}
